Open archaeology, open source?

### 2024-02-02 analysis/_06-network.qmd
Revise figures and tables
Removing superfluous code, consistent style, consistent captions and labels, etc.

```{r fig-graph-repo-user, eval=FALSE}
#| fig-cap: Interactive repository–user graph.
# TODO: is there a khroma scale that can accomodate this many groups?
g_repo_user <- oarch_graph_repo_user |>
  ggraph(layout = "fr", repulserad = 10, area = 10000, niter = 5000, grid = "nogrid") +
  geom_edge_link(aes(group = contrib, alpha = n_contrib)) +
  geom_point_interactive(aes(x, y, colour = group, shape=node_type, tooltip = node)) + 
  scale_edge_alpha(trans = "log", guide = guide_none()) +
  scale_colour_discrete(guide = guide_none()) +
  labs(edge_alpha = "Common contributors") +
  theme(panel.background = element_rect(fill = "white"))

@@

```{r fig-graph-repo-repo, eval=FALSE}
#| fig-cap: Interactive repository–repository graph. Repositories are connected by contributions by common users, with darker edges indicating more common contributors. Node colour indicates cluster membership according to the edge–betweenness method [@girvan2002]. Excludes isolate nodes.
# TODO: is there a khroma scale that can accomodate this many groups?
g_repo_repo <- oarch_graph_repo_repo |>
  ggraph(layout = "graphopt", mass = 100) +
  geom_edge_link(aes(alpha = n)) +
  geom_point_interactive(aes(x, y, colour = group, tooltip = repo)) + 
  scale_edge_alpha(trans = "log", guide = guide_none()) +
  scale_colour_discrete(guide = guide_none()) +
  labs(edge_alpha = "Common contributors") +
  theme(panel.background = element_rect(fill = "white"))

@@

```{r fig-graph-user-user, eval=FALSE}
#| fig-cap: Interactive user–user graph. Users are connected by contributions by common repositories, with darker edges indicating more common repositories. Node colour indicates cluster membership according to the edge–betweenness method [girvan2002]. Excludes isolate nodes.
# TODO: is there a khroma scale that can accomodate this many groups?
g_user_user <- oarch_graph_user_user |>
  filter(degree > 170) |>
  ggraph(layout = "fr", repulserad = 10, area = 100000, niter = 10000000, grid = "nogrid") +
  geom_edge_link(aes(alpha = n)) +
  geom_point_interactive(aes(x, y, tooltip = user)) + 
  scale_edge_alpha(trans = "log", guide = guide_none()) +
  scale_colour_discrete(guide = guide_none()) +
  labs(edge_alpha = "Common repositories") +
  theme(panel.background = element_rect(fill = "white"))

@@

<!-- Is there a network? How big is it? -->
Our data shows that there is a significant network of archaeologists collaborating on GitHub.
`r percent(graph_n_repos / graph_n_repos_total)` of repositories and `r percent(graph_n_users / graph_n_users_total)` of users in our dataset are connected to at least one other repository or user.
Of these, `r percent(max(subgraphs_n_repos$n) / graph_n_repos)` of repositories and `r percent(max(subgraphs_n_users$n) / graph_n_users)` of users belong to a single connected subgraph (@fig-graph-repo-repo and @fig-graph-user-user).

@@

```{r tbl-repo-betweenness}
#| tbl-cap: Most central repositories (weighted node betweenness) in the repository–repository network
oarch_graph_repo_repo |>
  mutate(betweenness = centrality_betweenness()) |>
  as_tibble() |>
  arrange(-betweenness) |>
  slice(1:10) |>
  left_join(oarch, c("repo" = "gh_repo")) |>
  mutate(n_commits = map_int(gh_commits, nrow)) |>
  unite(tags, tag1:tag5, sep = "; ", na.rm = TRUE) |>
  select(repo, betweenness, category, tags, n_commits) |>
  rename(repository = repo, commits = n_commits) |>
  rename_with(tools::toTitleCase) |>
  tt()
```

```{r fig-repo-betweenness-age-length}
#| fig-cap: Log-log correlation between repository centrality (weighted node betweenness), age (left) and length (right)
left <- oarch_graph_repo_repo |>
  as_tibble() |> 
  left_join(oarch, by = c("repo" = "gh_repo")) |>
  mutate(
    age = map_int(gh_commits, \(x) min(today() - date(x$datetime)))
  ) |>
  ggplot(aes(age, betweenness)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  geom_smooth(method = "glm") +
  labs(x = "Repository age (days)", y = "Centrality (betweenness)") +
  theme_oarch()

@@

```{r fig-repo-betweenness-category-platform}
#| fig-cap: Mean repository centrality (weighted node betweenness) by open-archaeo category (top) and platform (bottom)
top <- oarch_graph_repo_repo |>
  as_tibble() |> 
  left_join(oarch, by = c("repo" = "gh_repo")) |>
  group_by(category) |>
  summarise(betweenness = mean(betweenness)) |>
  mutate(category = fct_reorder(category, betweenness)) |>
  ggplot(aes(category, betweenness)) +
  geom_col() +
  scale_y_log10() +
  coord_flip() +
  labs(y = NULL, x = NULL) +
  theme_oarch()

bottom <- oarch_graph_repo_repo |>
  as_tibble() |> 
  left_join(oarch, by = c("repo" = "gh_repo")) |>
  drop_na(platform) |>
  group_by(platform) |>
  summarise(betweenness = mean(betweenness)) |>
  arrange(-betweenness) |>
  slice(1:7) |>
  mutate(platform = fct_reorder(platform, betweenness)) |>
  ggplot(aes(platform, betweenness)) +
  geom_col() +
  scale_y_log10() +
  coord_flip() +
  labs(y = "Centrality (weighted betweenness)", x = NULL) +
  theme_oarch()

@@

```{r fig-user-betweenness-contribs, eval=TRUE}
#| fig-cap: Log-log correlation between user centrality (weighted node betweenness) in the user-user graph and total number of contributions
oarch_graph_user_user |>
  as_tibble() |> 
  left_join(oarch_contributors, by = c("user" = "contributor")) |>
  ggplot(aes(betweenness, contribs)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  geom_smooth(method = "glm") +
  labs(x = "Centrality (betweenness)", y = "Total contributions") +
  theme_oarch()
```

>>>>

```{r fig-graph-repo-user}
#| fig-cap: Graph of open archaeology repositories and users connected by contributions. Darker edges indicate a great number of contributions. Node colour indicates cluster membership according to the edge–betweenness method [@girvan2002]. Excludes isolate nodes.
# TODO: is there a khroma scale that can accomodate this many groups?
g_repo_user <- oarch_graph_repo_user |>
  ggraph(layout = "fr", niter = 5000, grid = "nogrid") +
  geom_edge_link(aes(group = contrib, alpha = n_contrib)) +
  geom_point_interactive(aes(x, y, colour = group, shape=node_type, tooltip = node)) + 
  scale_edge_alpha(trans = "log", guide = guide_none()) +
  scale_colour_discrete(guide = guide_none()) +
  labs(shape = NULL) +
  theme(panel.background = element_rect(fill = "white"), legend.position = "bottom")

@@

```{r fig-graph-repo-repo}
#| fig-cap: Graph of open archaeology repositories connected by common contributors. Darker edges indicate a great number of common contributors. Node colour indicates cluster membership according to the edge–betweenness method [@girvan2002]. Excludes isolate nodes.
# TODO: is there a khroma scale that can accomodate this many groups?
g_repo_repo <- oarch_graph_repo_repo |>
  ggraph(layout = "graphopt", mass = 100) +
  geom_edge_link(aes(alpha = n)) +
  geom_point_interactive(aes(x, y, colour = group, tooltip = repo)) + 
  scale_edge_alpha(trans = "log", guide = guide_none()) +
  scale_colour_discrete(guide = guide_none()) +
  theme(panel.background = element_rect(fill = "white"))

@@

```{r fig-graph-user-user, eval=FALSE}
#| fig-cap: Graph of open archaeology users connected by contributions to common repositories. Darker edges indicate a great number of common repositories. Node colour indicates cluster membership according to the edge–betweenness method [@girvan2002]. Excludes isolate nodes.
# TODO: is there a khroma scale that can accomodate this many groups?
g_user_user <- oarch_graph_user_user |>
  filter(degree > 170) |>
  ggraph(layout = "fr", niter = 10000000, grid = "nogrid") +
  geom_edge_link(aes(alpha = n)) +
  geom_point_interactive(aes(x, y, tooltip = user)) + 
  scale_edge_alpha(trans = "log", guide = guide_none()) +
  scale_colour_discrete(guide = guide_none()) +
  theme(panel.background = element_rect(fill = "white"))

@@

<!-- Is there a network? How big is it? -->
Our data shows that there is a significant network of archaeologists collaborating on GitHub.
`r percent(graph_n_repos / graph_n_repos_total)` of repositories and `r percent(graph_n_users / graph_n_users_total)` of users in our dataset are connected to at least one other repository or user.
Of these, `r percent(max(subgraphs_n_repos$n) / graph_n_repos)` of repositories and `r percent(max(subgraphs_n_users$n) / graph_n_users)` of users belong to a single connected subgraph (@fig-graph-repo-user and @fig-graph-repo-repo).

@@

```{r tbl-repo-betweenness}
#| tbl-cap: Repositories ranked by centrality to the repository–repository network. Centrality is measured by node betweenness weighted by number of contributions.
oarch_graph_repo_repo |>
  as_tibble() |>
  arrange(-betweenness) |>
  slice(1:10) |>
  left_join(oarch, c("repo" = "gh_repo")) |>
  mutate(n_commits = map_int(gh_commits, nrow)) |>
  unite(tags, tag1:tag5, sep = "; ", na.rm = TRUE) |>
  transmute(rank = row_number(), repo, category, tags, n_commits) |>
  rename(repository = repo, commits = n_commits) |>
  rename_with(tools::toTitleCase) |>
  tt()
```

```{r fig-repo-betweenness-age-length}
#| fig-cap: Repository centrality by age (left) and length (right). Centrality is measured by node betweenness weighted by number of contributions.
left <- oarch_graph_repo_repo |>
  as_tibble() |> 
  left_join(oarch, by = c("repo" = "gh_repo")) |>
  mutate(
    age = map_int(gh_commits, \(x) min(today() - date(x$datetime)))
  ) |>
  ggplot(aes(age, betweenness)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  geom_smooth(method = "glm") +
  labs(x = "Repository age (days)", y = "Centrality") +
  theme_oarch()

@@

```{r fig-repo-betweenness-category-platform}
#| fig-cap: Mean repository centrality by category (top) and platform (bottom). Centrality is measured by node betweenness weighted by number of contributions.
top <- oarch_graph_repo_repo |>
  as_tibble() |> 
  left_join(oarch, by = c("repo" = "gh_repo")) |>
  group_by(category) |>
  summarise(betweenness = mean(betweenness)) |>
  mutate(category = fct_reorder(category, betweenness)) |>
  ggplot(aes(category, betweenness)) +
  geom_col() +
  scale_y_log10() +
  coord_flip() +
  labs(y = NULL, x = NULL) +
  theme_oarch()

bottom <- oarch_graph_repo_repo |>
  as_tibble() |> 
  left_join(oarch, by = c("repo" = "gh_repo")) |>
  drop_na(platform) |>
  group_by(platform) |>
  summarise(betweenness = mean(betweenness)) |>
  arrange(-betweenness) |>
  slice(1:7) |>
  mutate(platform = fct_reorder(platform, betweenness)) |>
  ggplot(aes(platform, betweenness)) +
  geom_col() +
  scale_y_log10(labels = NULL) +
  coord_flip() +
  labs(y = "Centrality →", x = NULL) +
  theme_oarch()

@@

```{r fig-user-betweenness-contribs}
#| fig-cap: User centrality by total number of contributions. Centrality is measured by node betweenness weighted by number of contributions.
oarch_graph_user_user |>
  as_tibble() |> 
  left_join(oarch_contributors, by = c("user" = "contributor")) |>
  ggplot(aes(contribs, betweenness)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  geom_smooth(method = "glm") +
  labs(x = "Total contributions", y = "Centrality") +
  theme_oarch()
```

### 2024-02-02 analysis/openarchaeo_collaboration.qmd
Revise figures and tables
Removing superfluous code, consistent style, consistent captions and labels, etc.

library("english")
library("fs")
library("geomtextpath")
library("ggiraph")
library("ggraph")
library("gt")
library("here")
library("khroma")
library("lubridate")
library("patchwork")
library("scales")
library("tidygraph")
library("tidyverse")
```

>>>>

library("english")
library("fs")
library("ggdist")
library("ggiraph")
library("ggraph")
library("here")
library("khroma")
library("lubridate")
library("patchwork")
library("rgexf")
library("scales")
library("tidygraph")
library("tidyverse")
library("tinytable")
```
